---
title: "Phone Usage Tracking Report"
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}
h1 {
  color: #fafafc;
  font-family: 'Futura', sans-serif;
  font-weight: bold;
  background-color: #0f0f42;
  padding: 20px;
}

h2 {
  color: #fafafc;
  font-family: 'PT Sans', sans-serif;
  background-color: #0f0f42;
  font-weight: bold;
  padding: 20px;
  border-bottom: 3px solid #4f87db;
}

body {
  font-family: 'PT Sans', sans-serif;
  background: linear-gradient(135deg, #b3d1ff, #d1deeb);
  font-size: 18px;
  line-height: 1.6;
  letter-spacing: 0.5px;
  color: #333;
}

a {
  color: #0f0f42;
  text-decoration: none;
  transition: color 0.3s ease;
}


img {
  border: 5px solid #0f0f42;
  border-radius: 10px;
  margin: 20px 0;
}


ul {
  list-style-type: square;
}
```


# Overview of Phone Usage Data

We explain the trend of mobile phone usage over time here in this report.

The data gathered from it have information on:

- The amount of hours spent using one's phones.

- The primary thing done on their phones.

- At what time of day people used their phones the most.

- Phone usage Locations by Time of Day

For the purposes of analysis, this report will look at the data through major summary data and charts.

## Summary of key stats
We start with the most essential information regarding entertainment and phone use.


```{r}

library(tidyverse)

logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSnGZftiMapz_Zg6q_MLISNyoJ3PQ2kFAnSDlLvuXmJQuxxTAMDmfvt9W-rILHm2Ez1NyIkQMsqTQZy/pub?gid=296253976&single=true&output=csv")


latest_data <- logged_data %>%
  rename(
    time_spent = 2,
    usage_purpose = 3,
    phone_usage_location = 4,
    most_phone_usage_time_of_day = 5,
    entertainment_time = 6
  )


# Create new columns for each option in the "usage_purpose" question
latest_data <- latest_data %>%
  mutate(
    messaging = str_detect(usage_purpose, "Messaging"),
    social_media = str_detect(usage_purpose, "Social media"),
    entertainment = str_detect(usage_purpose, "Entertainment"),
    work_study = str_detect(usage_purpose, "Work/Study"),
    stalking_workout = str_detect(usage_purpose, "Stalking both my beautiful ex situationships relentlessly|Workout Tracking")
  )


# ===============================
# Summary Statistics
# ===============================


# Identify minimum, maximum, and mean values for phone usage time
min_time <- min(latest_data$time_spent, na.rm = TRUE)
max_time <- max(latest_data$time_spent, na.rm = TRUE)
mean_time <- mean(latest_data$time_spent, na.rm = TRUE)

# Identify minimum, maximum, and mean values for entertainment time
min_entertainment <- min(latest_data$entertainment_time, na.rm = TRUE)
max_entertainment <- max(latest_data$entertainment_time, na.rm = TRUE)
mean_entertainment <- mean(latest_data$entertainment_time, na.rm = TRUE)


# Dynamic comments based on the summary values
paste("The minimum time spent on the phone in a day was", min_time, "minutes while the maximum was", max_time, "minutes. On average, people spent", round(mean_time, 1), "minutes on their phones.")
paste("For entertainment, the minimum time spent was", min_entertainment, "minutes, and the average time spent on entertainment was", round(mean_entertainment, 1), "minutes.")

```


## Usage Purpose Breakdown
The breakdown of phone usage purpose is depicted in the following bar chart. It informs us what individuals do with their phones most of the time.


```{r}
# 1. Bar Chart: Frequency of Usage Purpose
usage_count <- latest_data %>%
  summarise(
    Messaging = sum(messaging, na.rm = TRUE),
    Social_Media = sum(social_media, na.rm = TRUE),
    Entertainment = sum(entertainment, na.rm = TRUE),
    Work_Study = sum(work_study, na.rm = TRUE),
    Other = sum(stalking_workout, na.rm = TRUE)
  )

# Convert counts to long format for plotting
usage_count_long <- usage_count %>%
  pivot_longer(cols = everything(), names_to = "usage_type", values_to = "count")

# Plot bar chart for usage purposes
ggplot(usage_count_long, aes(x = usage_type, y = count, fill = usage_type)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Frequency of Phone Usage Purposes",
    x = "Usage Type",
    y = "Count"
  )

```

## Time of Day When People Use Their Phones the Most
We then look at the times of day when people are most likely to use their phones. The most common usage times are marked in the graph below.

``` {r}
# 2. Bar Graph: Showing Most Common Phone Usage Time of Day

# Define Order for Time Categories (1 = Morning, 5 = Late Night)

time_labels <- c("1" = "Morning", "2" = "Late Morning", "3" = "Afternoon", 
                 "4" = "Evening", "5" = "Late Night")


latest_data$most_phone_usage_time_of_day <- factor(latest_data$most_phone_usage_time_of_day, 
                                                   levels = c("1", "2", "3", "4", "5"))


# All Time Periods Appear in the Data Frame

time_of_day_counts <- latest_data %>%
  count(most_phone_usage_time_of_day) %>%
  complete(most_phone_usage_time_of_day = factor(c("1", "2", "3", "4", "5"), levels = c("1", "2", "3", "4", "5")), 
           fill = list(n = 0))  # Fill missing values with 0


ggplot(time_of_day_counts, aes(x = most_phone_usage_time_of_day, y = n, fill = most_phone_usage_time_of_day)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(labels = time_labels) +  
  theme_minimal() +
  labs(
    title = "Most Common Phone Usage Time of Day",
    x = "Time of Day",
    y = "Frequency"
  ) +
  theme(legend.position = "none")
```


## Phone Usage Locations by Time of Day
The stacked bar chart shows that phone usage is highest in the evening and late night, primarily at home, while usage during commuting, in class, and on lunch breaks is less frequent throughout the day.






## Conclusion
Various conclusions have been drawn from the analysis:

- In average, individuals use their phones mostly for social media and/or messaging for a significant amount of time.

- The most common time of day phone was used is late night.

- When the phone was used most was mostly at home. 

The dynamic feature of the report ensures that the charts and summary will automatically refresh every time additional information is collected via the survey.



