---
title: "Phone Usage Tracking Report"
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}
h1 {
  color: #fafafc;
  font-family: 'Futura', sans-serif;
  font-weight: bold;
  background-color: #0f0f42;
  padding: 20px;
}

h2 {
  color: #fafafc;
  font-family: 'PT Sans', sans-serif;
  background-color: #0f0f42;
  font-weight: bold;
  padding: 20px;
  border-bottom: 3px solid #4f87db;
}

body {
  font-family: 'PT Sans', sans-serif;
  background: linear-gradient(135deg, #b3d1ff, #d1deeb);
  font-size: 18px;
  line-height: 1.6;
  letter-spacing: 0.5px;
  color: #333;
}

a {
  color: #0f0f42;
  text-decoration: none;
  transition: color 0.3s ease;
}


img {
  border: 5px solid #0f0f42;
  border-radius: 10px;
  margin: 20px 0;
}


ul {
  list-style-type: square;
}
```


## Overview of Phone Usage Data

In this report, we track the usage of mobile phones over a period of time. 

The data collected includes information on:


- How much time people spend on their phones.


- What they mainly use their phones for.


- The most frequent times of day people use their phones.


This report will analyze the data using key summary values and visualizations to provide insights.

Summary of Key Statistics
We begin by looking at the key statistics for time spent on phones and the time spent on entertainment.

```{r}

library(tidyverse)

logged_data <- read_csv("/Users/keerurangu/Downloads/Phone Usage Tracking Survey (Responses) - Form responses 1.csv")

latest_data <- logged_data %>%
  rename(
    time_spent = 2,
    usage_purpose = 3,
    phone_usage_location = 4,
    most_phone_usage_time_of_day = 5,
    entertainment_time = 6
  )


# Create new columns for each option in the "usage_purpose" question
latest_data <- latest_data %>%
  mutate(
    messaging = str_detect(usage_purpose, "Messaging"),
    social_media = str_detect(usage_purpose, "Social media"),
    entertainment = str_detect(usage_purpose, "Entertainment"),
    work_study = str_detect(usage_purpose, "Work/Study"),
    stalking_workout = str_detect(usage_purpose, "Stalking both my beautiful ex situationships relentlessly|Workout Tracking")
  )



# Identify minimum, maximum, and mean values for phone usage time
min_time <- min(latest_data$time_spent, na.rm = TRUE)
max_time <- max(latest_data$time_spent, na.rm = TRUE)
mean_time <- mean(latest_data$time_spent, na.rm = TRUE)

# Identify minimum, maximum, and mean values for entertainment time
min_entertainment <- min(latest_data$entertainment_time, na.rm = TRUE)
max_entertainment <- max(latest_data$entertainment_time, na.rm = TRUE)
mean_entertainment <- mean(latest_data$entertainment_time, na.rm = TRUE)


# Dynamic comments based on the summary values
paste("The minimum time spent on the phone in a day was", min_time, "minutes, while the maximum was", max_time, "minutes. On average, people spent", round(mean_time, 1), "minutes on their phones.")
paste("For entertainment, the minimum time spent was", min_entertainment, "minutes, and the average time spent on entertainment was", round(mean_entertainment, 1), "minutes.")

```


## Usage Purpose Breakdown
Below is a bar chart that shows the distribution of phone usage purposes. It gives us an idea of what people primarily use their phones for.

```{r}
library(tidyverse)

logged_data <- read_csv("/Users/keerurangu/Downloads/Phone Usage Tracking Survey (Responses) - Form responses 1.csv")

latest_data <- logged_data %>%
  rename(
    time_spent = 2,
    usage_purpose = 3,
    phone_usage_location = 4,
    most_phone_usage_time_of_day = 5,
    entertainment_time = 6
  )


# Create new columns for each option in the "usage_purpose" question
latest_data <- latest_data %>%
  mutate(
    messaging = str_detect(usage_purpose, "Messaging"),
    social_media = str_detect(usage_purpose, "Social media"),
    entertainment = str_detect(usage_purpose, "Entertainment"),
    work_study = str_detect(usage_purpose, "Work/Study"),
    stalking_workout = str_detect(usage_purpose, "Stalking both my beautiful ex situationships relentlessly|Workout Tracking")
  )

# 1. Bar Chart: Frequency of Usage Purpose
usage_count <- latest_data %>%
  summarise(
    Messaging = sum(messaging, na.rm = TRUE),
    Social_Media = sum(social_media, na.rm = TRUE),
    Entertainment = sum(entertainment, na.rm = TRUE),
    Work_Study = sum(work_study, na.rm = TRUE),
    Other = sum(stalking_workout, na.rm = TRUE)
  )

# Convert counts to long format for plotting
usage_count_long <- usage_count %>%
  pivot_longer(cols = everything(), names_to = "usage_type", values_to = "count")

# Plot bar chart for usage purposes
ggplot(usage_count_long, aes(x = usage_type, y = count, fill = usage_type)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Frequency of Phone Usage Purposes",
    x = "Usage Type",
    y = "Count"
  )

```

## Most Common Phone Usage Time of Day
Next, we explore the most common times of day that people use their phones. The following bar chart highlights the most frequent usage times.

``` {r}
library(tidyverse)

logged_data <- read_csv("/Users/keerurangu/Downloads/Phone Usage Tracking Survey (Responses) - Form responses 1.csv")

latest_data <- logged_data %>%
  rename(
    time_spent = 2,
    usage_purpose = 3,
    phone_usage_location = 4,
    most_phone_usage_time_of_day = 5,
    entertainment_time = 6
  )


# Create new columns for each option in the "usage_purpose" question
latest_data <- latest_data %>%
  mutate(
    messaging = str_detect(usage_purpose, "Messaging"),
    social_media = str_detect(usage_purpose, "Social media"),
    entertainment = str_detect(usage_purpose, "Entertainment"),
    work_study = str_detect(usage_purpose, "Work/Study"),
    stalking_workout = str_detect(usage_purpose, "Stalking both my beautiful ex situationships relentlessly|Workout Tracking")
  )


# 2. Bar Graph: Showing Most Common Phone Usage Time of Day

# Define Order for Time Categories (1 = Morning, 5 = Late Night)

time_labels <- c("1" = "Morning", "2" = "Late Morning", "3" = "Afternoon", 
                 "4" = "Evening", "5" = "Late Night")


latest_data$most_phone_usage_time_of_day <- factor(latest_data$most_phone_usage_time_of_day, 
                                                   levels = c("1", "2", "3", "4", "5"))


# Ensure All Time Periods Appear in the Data Frame

time_of_day_counts <- latest_data %>%
  count(most_phone_usage_time_of_day) %>%
  complete(most_phone_usage_time_of_day = factor(c("1", "2", "3", "4", "5"), levels = c("1", "2", "3", "4", "5")), 
           fill = list(n = 0))  # Fill missing values with 0


ggplot(time_of_day_counts, aes(x = most_phone_usage_time_of_day, y = n, fill = most_phone_usage_time_of_day)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(labels = time_labels) +  
  theme_minimal() +
  labs(
    title = "Most Common Phone Usage Time of Day",
    x = "Time of Day",
    y = "Number of People"
  ) +
  theme(legend.position = "none")
```

## Conclusion
From the analysis, we can draw several insights:

- People on average spend a significant amount of time on their phones for social media and entertainment.

- The peak usage time of day tends to be in the Evening, followed by the Afternoon.


The dynamic nature of the report ensures that as more data is collected through the survey, the charts and summaries will update automatically.




